---
title: "Reproducible_Research_Proj1"
author: "JESTO"
date: "11/13/2020"
output: html_document
    keep_md: true
---

Loading ggplot2, dplyr package and data set.  Manipulating data to make it workable
====================================================================================

```{r, echo=TRUE}
library(ggplot2)
library(dplyr)
activity <- read.csv("./activity.csv", sep = ",", header = TRUE)
head(activity)
names(activity)
str(activity)
```
Convert date format to date
```{r, echo=TRUE}
activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
activity <- mutate(activity, weekday=weekdays(date))
head(activity)
names(activity)
str(activity)
summary(activity)
```

Total Steps Histogram total number of steps taken per day. Output mean, and median 
==================================================================================

```{r, echo=TRUE}
TotalStepsDay <- aggregate(steps ~ date, activity, FUN = sum, na.rm=TRUE)
head(TotalStepsDay)
TotalStepsPlot <- ggplot(TotalStepsDay, aes(steps))
TotalStepsPlot + geom_histogram(bins=15, fill="red", color="black") + labs(title="Histogram Total Number of Steps per Day", x="Total Number of Steps", y="Frequency")
```

Mean of steps per day
```{r, echo=TRUE}
mean(TotalStepsDay$steps, na.rm = TRUE)
```

Median of steps per day
```{r, echo=TRUE}
median(TotalStepsDay$steps, na.rm = TRUE)
```


Average Daily Activity (Steps) per Interval
==============================================================

```{r, echo=TRUE}
AverageActivityInterval <- aggregate(steps ~ interval, activity, FUN = mean, na.rm=TRUE)
names(AverageActivityInterval) <- c("interval", "mean.steps")
head(AverageActivityInterval)
AverageActivityPlot <- ggplot(AverageActivityInterval, aes(x= interval, y = mean.steps))
AverageActivityPlot + geom_line() + labs(title="Average Steps per Interval", x="Interval", y="Average Steps", color="blue")
```

Which 5 minute interval has the highest average activity?
```{r, echo=TRUE}
AverageActivityInterval[which.max(AverageActivityInterval$mean.steps), ]$interval

```




## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
