---
title: "Revised_Reproducible_Research_Proj1"
author: "JESTO"
date: "11/15/2020"
output: html_document
---

Remove Date Warnings - Global Option
===================================================================================

```{r}
knitr::opts_chunk$set(warning=FALSE)
```


Loading ggplot2, dplyr package and data set.  Manipulating data to make it workable
====================================================================================

```{r, echo=TRUE}
library(ggplot2)
library(dplyr)
activity <- read.csv("./activity.csv", sep = ",", header = TRUE)
head(activity)
names(activity)
str(activity)
```

*Convert date format to date*
```{r, echo=TRUE}
activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
## activity <- transform(activity, date = strptime(date, "%Y-%m-%d"))
activity <- mutate(activity, weekday=weekdays(date))
head(activity)
names(activity)
str(activity)
summary(activity)
```

Total Steps Histogram total number of steps taken per day. Output mean, and median 
==================================================================================

```{r, echo=TRUE}
TotalStepsDay <- aggregate(steps ~ date, activity, FUN = sum, na.rm=TRUE)
head(TotalStepsDay)
TotalStepsPlot <- ggplot(TotalStepsDay, aes(steps))
TotalStepsPlot + geom_histogram(bins=15, fill="red", color="black") + labs(title="Histogram Total Number of Steps per Day", x="Total Number of Steps", y="Frequency")
```

*Mean of steps per day*
```{r, echo=TRUE}
mean(TotalStepsDay$steps, na.rm = TRUE)
```

*Median of steps per day*
```{r, echo=TRUE}
median(TotalStepsDay$steps, na.rm = TRUE)
```


Average Daily Activity (Steps) per Interval
==============================================================

```{r, echo=TRUE}
AverageInterval <- aggregate(steps ~ interval, activity, FUN = mean, na.rm=TRUE)
names(AverageInterval) <- c("interval", "mean_steps")
head(AverageInterval)
AverageActivityPlot <- ggplot(AverageInterval, aes(x= interval, y = mean_steps))
AverageActivityPlot + geom_line() + labs(title="Average Steps per Interval", x="Interval", y="Average Steps", color="blue")
```

*Which 5 minute interval has the highest average activity?*
```{r, echo=TRUE}
AverageInterval[which.max(AverageInterval$mean_steps), ]$interval

```


Inputting Missing Values
======================================================

*Determine number of missing values*
```{r, echo=TRUE}
TotalNAs <- sum(is.na(activity$steps))
TotalNAs
```

**Substitute missing values with mean per interval**

*First I create a data frame matching intervals*

```{r, echo=TRUE}
SubstituteSteps <- AverageInterval$mean_steps[match(activity$interval, AverageInterval$interval)]
head(SubstituteSteps)
```

*Second Create a new data set with NA substituted by mean value*
```{r, echo=TRUE}
ModActivity <- transform(activity, steps=ifelse(is.na(activity$steps), SubstituteSteps, activity$steps))
head(ModActivity)
TotalNAs1 <- sum(is.na(ModActivity))
TotalNAs1
```

*Third create a histogram for the modified dataset*
```{r, echo=TRUE}
ModTotalStepsDay <- aggregate(steps ~ date, ModActivity, FUN = sum, na.rm=TRUE)
head(ModTotalStepsDay)
ModTotalStepsPlot <- ggplot(ModTotalStepsDay, aes(steps))
ModTotalStepsPlot + geom_histogram(bins=15, fill="green", color="black") + labs(title="Histogram Total Number of Steps per Day", x="Total Number of Steps", y="Frequency")
```

*Modified mean of steps per day*
```{r, echo=TRUE}
mean(ModTotalStepsDay$steps, na.rm = TRUE)
```

*Modified median of steps per day*
```{r, echo=TRUE}
median(ModTotalStepsDay$steps, na.rm = TRUE)
```

Difference in activity between weekdays and weekend
========================================================

*Create a new variable name day_type*
```{r, echo=TRUE}
DayModActivity <- mutate(activity, day_type = ifelse((weekday == "Saturday" | weekday == "Sunday"), "weekend", "weekday"))
head(DayModActivity)
```

*Aggregate steps by interval and day typ
```{r, echo=TRUE}
AvgDayModActivity <- aggregate(steps ~ interval + day_type, DayModActivity, FUN=mean, na.rm=TRUE )
head(AvgDayModActivity)
```


*Create a panel plot showing activity on weekdays against activity on weekends*
```{r, echo=TRUE}
DayModActivityPlot <- ggplot(AvgDayModActivity, aes(x=interval, y=steps, color=day_type))
DayModActivityPlot + geom_line() + facet_grid(day_type~.) + labs(title = "Average Activity depending on  day of the week", x = "Interval", y = "Average Number of Steps")
        
```
